The 2018 heatwave and its implications on ozone induced damage on vegetation in a subarctic climate

### 2020-04-29 VegOzone.tex
Include ozone reconstruction plot again.
Added plots and analyse for ozone reconstruction ion 2018.

An unusually weak and northward shifted jet stream allowed for a persistent high pressure system above northern Europe, including Fennoscandia which blocked the pressure systems for several consecutive months. In the periode May -- July, southern Norway had temperatures $4\,\unit{^\circ C}$ above normal. Whgile southern Norway had only about $60\,\unit{\%}$ of normal precipitation, northern Norway as a whle had close to normalprecipitation, but with local variations.

@@

Elevated ozone in Europe was liekly due to several other factors as well, like in the 2003 drought period, such as biogenic VOC emissions, reduced dry deposition, reduced clouds and increased solar radiation, photochemical reactions, and stagnant weather conditions \citep{JGR:Solberg2018}\\

@@

The density distribution for Svanvik and the respective \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, mean}}, and \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, max}} (Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_spline}b) display a similar pattern as the northern Fennoscandic climatology. The spring peak apparently occurs slightly earlier (in the beginning of April). As Svanvik is located at lower altitude than the other sites, the earlier ozone peak may point to an interplay with ozone uptake by vegetation. {\bf TODO: The data from SeNorge seems to point to a later onset of greening, than what I interfered from OpenIFS data. This means that the beginning of the growing season is not a good measure for the determination of the ozone spring peak. This does not exclude the early start of pine forest photosynthesis as a driver. Is there a way to confirm this? Actual data for the three sites? Please correct if wrong! Ane? Aud?} The average ozone concentration at Svanvik is lower than at the other sites, as already mentioned, which is in league with a decrease in tropospheric ozone with decreasing altitude. In July--September, ozone is occasionally almost completely depleted during night time.

@@

\section{DO3SE modeling}
\label{sec:do3se}
To investigate which of the environmental factors lead to the observed ozone damage on clover and tobacco {\bf TODO: Ane confirm!} at Svanhovd, we have modeled the ozone uptake using the DO3SE model.

@@

\begin{figure}[t]
  \includegraphics[width=8.3cm]{ozone_climatology_fenoscandic_obs_test-Svanvik}
  \caption{Student's t-test assuming same sample uncertainty in both, climatology and 2018/19 observations at Svanhovd with respect to derived climatologies for (a) Northern Fennoscandia; (b) Svanvik. \chem{\Delta[O_3]} in Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals}a) are significantly different from zero-hypothesis on the $2\,\sigma$,  $1\,\sigma$ level, respectively. \chem{\Delta[O_3]} in Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals}b) are not significantly different from zero-hypothesis.}
  \label{fig:ozone_climatology_fenoscandic_obs_test-Svanvik}
\end{figure}
\clearpage

@@

%\begin{figure}[t]
%  \includegraphics[width=8.3cm]{ozone_obs_2018_07}
%  \caption{Estimate of missing observation \chem{[O_3]} data at Svanhovd in July 2018 based on derived hourly climatologies of northern Fennoscandia, Svanvik, and Copernicus regional model reanalysis ensemble mean. (a) Time series; (b) Student's t-test.}
%  \label{fig:ozone_obs_2018_rrea}
%\end{figure}

>>>>

An unusually weak and northward shifted jet stream allowed for a persistent high pressure system above northern Europe, including Fennoscandia which blocked the pressure systems for several consecutive months. In the period May -- July, southern Norway had temperatures $4\,\unit{^\circ C}$ above normal. While southern Norway had only about $60\,\unit{\%}$ of normal precipitation, northern Norway as a whole had close to normal precipitation, but with local variations.

@@

Elevated ozone in Europe was likely due to several other factors as well, like in the 2003 drought period, such as biogenic VOC emissions, reduced dry deposition, reduced clouds and increased solar radiation, photochemical reactions, and stagnant weather conditions \citep{JGR:Solberg2018}.\\

@@

The density distribution for Svanvik and the respective \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, mean}}, and \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, max}} (Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_spline}b) display a similar pattern as the northern Fennoscandic climatology. The spring peak apparently occurs slightly earlier (in the beginning of April). As Svanvik is located at lower altitude than the other sites, the earlier ozone peak may point to an interplay with ozone uptake by vegetation. {\bf TODO: The data from SeNorge seems to point to a later onset of greening, than what I interfered from OpenIFS data. This means that the beginning of the growing season is not a good measure for the termination of the ozone spring peak. This does not exclude the early start of pine forest photosynthesis as a driver. Coniferous forest starts to photosynthesis earlier than what the growing season begin would mark. I remember Håkan or Per Erik(?) talking about the \chem{CO_2} uptake and phenology of coniferous trees - but still have not found a related publication? Is there a way to confirm this? Actual data for the three sites? Please correct if wrong! Ane? Aud?} The average ozone concentration at Svanvik is lower than at the other sites, as already mentioned, which is in league with a decrease in tropospheric ozone with decreasing altitude. In July--September, ozone is occasionally almost completely depleted during night time.

@@

\section{DO3SE modeling}
\label{sec:do3se}
To investigate which of the environmental factors lead to the observed ozone damage on clover and tobacco in 2018 in contrast to 2019, when no such damage was found on the clovers but only on the sensitive tobacco cultivar, we model the ozone uptake using the DO3SE model.

\subsection{Reconstruction of missing ozone values in 2018}
\label{subsec:ozone_reco}
As we have shown, the available reanalysis products have their limitations. Although the quality of the Copernicus Regional Model Reanalysis is fairly good, we additionally deem it necessary to reconstruct the missing data in 2018 for the purpose of DO3SE modeling. In Appendix Fig.~\ref{fig:ozone_reconstruction_2018_07}, the procedure for a probable reconstruction is shown. First, we derive hourly climatologies for northern Fennoscandia \chem{\left<[O_3]\right>_{hourly}} and Svanvik \chem{\left<[O_3]\right>_{hourly}^{Svanvik}} (Appendix Fig.~\ref{fig:ozone_reconstruction_2018_07}a). Then we compute temporal correlations between Svanvik and all other stations with respect to a temporal shift (Appendix Fig.~\ref{fig:time_lag_correlation}) to estimate a probable time lag between the stations (e.g., $3\,\unit{h}$ for Esrange and Pallas). Because of the high correlation with Svanvik observations, we have chosen Pallas as reference station for the reconstruction. We correct the Fennoscandic climatology for the found time lag\chem{\left<[O_3]\right>_{hourly,\,t-3}} and use it to derive a mapping function:
\begin{equation}
  X = \frac{\chem{\left<[O_3]\right>_{hourly}^{Svanvik}}}{\chem{\left<[O_3]\right>_{hourly,\,t-3}}}.
  \label{eq:mapping}
\end{equation}
Performing a time lag correction basically means to shift the respective time series back in time by $3\,\unit{h}$.
As an intermediate step, hourly anomalies are computed
\begin{equation}
  \chem{\Delta[O_3]_{hourly}^{Pallas}} = \chem{[O_3]_{hourly}^{Pallas}}-\chem{\left<[O_3]\right>_{hourly}}.
\end{equation}
In Appendix Fig.~\ref{fig:ozone_reconstruction_2018_07}b), hourly ozone concentration anomalies are shown for Pallas as well as for Esrange and Svanvik. We also correct ozone anomalies for Pallas for the time lag \chem{\Delta[O_3]_{hourly,\,t-3}^{Pallas}} and use the mapping function in Eq.~(\ref{eq:mapping}) to derive reconstructed anomalies for the missing values:
\begin{equation}
    \chem{\Delta [O_3]_{hourly}^{Svanvik,\,reco}} = \chem{\Delta[O_3]_{hourly,\,t-3}^{Pallas}} \cdot X
\end{equation}
Finally, we add the bias corrected Svanvik ozone climatology and derive a reconstructed ozone concentration time series \chem{[O_3]_{hourly}^{Svanvik,\,reco}}. In Appendix Fig.~\ref{fig:ozone_reconstruction_2018_07}c), the result is shown together with the observed data and the Copernicus Regional Model Reanalysis at the nearest grid point for comparison.

\subsection{Model description and input data}

{\bf TODO: Some general words about DO3SE model?}

\subsection{POD and ozone impact}

@@

\begin{figure}[t]
  \includegraphics[width=8.3cm]{ozone_climatology_fenoscandic_obs_test-Svanvik}
  \caption{Student's t-test assuming same sample uncertainty in both, climatology and 2018/19 observations at Svanhovd with respect to derived climatologies for (a) Northern Fennoscandia; (b) Svanvik. \chem{\Delta[O_3]} in Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals}a) are significantly different from zero-hypothesis on the $2\,\sigma$,  $1\,\sigma$ level, respectively. \chem{\Delta[O_3]} in Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals}b) are not significantly different from zero-hypothesis.}
  \label{fig:ozone_climatology_fenoscandic_obs_test-Svanvik}
\end{figure}

\begin{figure}[t]
  \includegraphics[width=8.3cm]{ozone_reconstruction_2018_07}
  \caption{Reconstruction of missing \chem{[O_3]} observation data at Svanvik in July 2018 based on derived hourly climatologies of northern Fennoscandia, Svanvik, and data Pallas. (a) Time series; (b) anomalies; (c) reconstruction.}
  \label{fig:ozone_reconstruction_2018_07}
\end{figure}

\begin{figure}[t]
  \includegraphics[width=8.3cm]{ozone_observation_timelag_svanvik}
  \caption{Temporal correlation of \chem{[O_3]} data at Svanvik with other ozone monitoring stations in northern Fennoscandia. A negative lag means that Svanvik lags behind, while a positive lag mean the other station lags behind. The highest correlation with Pallas/Esrange is found at a time lag of $3\,\unit{h}$, for Jergul/Karasjok at $1\,\unit{h}$, and for Janiskoski at $-20\,\unit{h}$.}
  \label{fig:time_lag_correlation}
\end{figure}

\clearpage

@@
